---
title: "Untitled"
author: "Zanne"
date: "7/22/2019"
output:
  word_document: default
  html_document: default
editor_options:
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Outline of six Month Morphology Paper

```{r cars, echo=FALSE, include=FALSE, warning='hide'}

t=read.csv('MorphData29 May 2020.csv')
ll = read.csv('LegLengths.csv')
es = read.csv('ElasticData25 Nov 2020.csv')


es$tendonEGpa = es$tendonE/1e9
es$KE = 0.5*es$mass*es$maxV^2
es$workPKg = es$work/(2*es$sumExtensors/1000)

es$legLength = es$femur+es$tibia+es$tmt+es$toe
es$femurN = es$femur/es$legLength
es$tibiaN = es$tibia/es$legLength
es$tmtN = es$tmt/es$legLength
es$toeN = es$toe/es$legLength

ed = es[es$group=='E' |es$group=='D',]
ed$group = droplevels(ed$group)


```

```{r Functions, echo=FALSE, include=FALSE, warning='hide'}
require('lme4')
require('stringr')
require('doBy')
require('plyr')
require('extrafont')
require('MHTdiscrete')
require('psych')
require('lsr')
require('MASS')
loadfonts(device='win')

windowsFonts(
  A=windowsFont("Garamond"),
  B=windowsFont("Aerial"),
  C=windowsFont("Times New Roman")
)


lmp <- function (modelobject) {
  if (class(modelobject) != "lm") stop("Not an object of class 'lm' ")
  f <- summary(modelobject)$fstatistic
  p <- pf(f[1],f[2],f[3],lower.tail=F)
  attributes(p) <- NULL
  return(p)
}

boxPoint = function(var,fact,titletxt, cols, groupNames){
    resp=var
 treat=fact
 par(family='Garamond')
 if(missing(cols)) {cols = 'grey'}
bTest =  bartlett.test(resp~treat)
if (bTest$p.value<0.05) {nonparametric = 1}
else {nonparametric = 0}

if (nonparametric)  {
  Test = kruskal.test(resp~treat)
  pVal = Test$p.value}
else {
  if (nlevels(treat)==2){
    Test= t.test(resp~treat)
    pVal = Test$p.value}
  else {pVal = anova(lm(mass~group,data=d))$"Pr(>F)"[1]}
}
if ((missing(groupNames))){
  groupNames = levels(var)
}
if ((missing(titletxt))){
  titletxt = 'variable'
}

titletxt = str_c(titletxt,' p:', format(pVal,digits=2))
boxplot(resp~treat, main=titletxt,las=2,col=cols, cex.axis=1.2, names = groupNames)
stripchart(resp~treat, vertical=TRUE, add=TRUE, pch=16,col='red',method='jitter',cex=1)
}



plotStats = function(dependent,independent,dta,fctor, titletxt, cols){
  x=unlist(dta[names(dta)==independent])
  y=unlist(dta[names(dta)==dependent])
 group = unlist(dta[names(dta)==fctor])
 lmFit = lm(x~y)
pValFit = summary(lmFit)
if ((missing(titletxt))){titletxt = str_c(dependent,' vs ',independent)}
if ((missing(cols))){cols = c('white','grey50')}
titletxt = str_c(titletxt,' p:',format(pValFit$coefficients[2,4],digits=2))
plot(y~x,main =titletxt , pch=21,bg=cols[as.numeric(group)],col='black', cex=1.2, xlab = independent, ylab = dependent)
legend('bottomright',levels(group),pch=21,pt.bg=cols,cex=1.2,col='black',inset=c(0,1), xpd=TRUE, horiz=TRUE, bty="n")
}


tableFormatByGroup = function(var,fctor,dta){
  y=unlist(dta[var])
  group =unlist(dta[fctor])
  meanVar = aggregate(y~group,data=dta,function(x) mean(x,na.rm=TRUE),na.action=na.pass)
  sdVar = aggregate(y~group,data=dta,function(x) sd(x,na.rm=TRUE),na.action=na.pass)
  if (sdVar$y[1] ==0 |sdVar$y[2] ==0){
    nonparametric=0
    pVal = 1
  } else {
    bTest =  bartlett.test(y~group,data=dta)
    nonparametric = ifelse(bTest$p.value<0.05,1,0)
    
    if (nonparametric)  {
      Test = kruskal.test(y~group,data=dta)
      pVal = Test$p.value
    } else {
      if (nlevels(group)==2){
        Test= t.test(y~group)
        pVal = Test$p.value
      } else {pVal = anova(lm(y~group,data=dta))$"Pr(>F)"[1]}
    }
  }
  cohd = cohensD(y~group)
  varString = ' '
for (i in 1:nlevels(group)){varString[i] = str_c(format(meanVar[i,2],digits=2),'+-',format(sdVar[i,2],digits=2))}
varString[i+1] = ifelse(nonparametric,str_c(format(pVal,digits=2),'*'),str_c(format(pVal,digits=2)))

varString[i+2]= format(cohd, digits=2)
return(varString)
}




```

## Questions  a) Reductionist Level

1. Do components of MTU's change in response to changes in functional demand?  Specifically, Do individuals with the same functional demands adapt their MTU components is a systematic way?



```{r tableFunctional1, echo=FALSE, results='asis', fig.width=8, warning=FALSE, message=FALSE, fig.height=7}

dta=ed
require('pander')
panderOptions('round',1)
panderOptions('table.style','rmarkdown')
panderOptions("table.split.table", Inf) 
panderOptions('keep.trailing.zeros',TRUE)
library(plyr)
#functional data
totalsEd = count (ed$group)
totalsEd[3,2] = ' '
totalsEd[4,2] = ' '
massString = tableFormatByGroup('mass','group',ed)
bDensityString = tableFormatByGroup('BmdGcm2','group',ed)
perFatString =  tableFormatByGroup('PerFat','group',ed)
avMaString =  tableFormatByGroup('avMa','group',ed)
tendonKString = tableFormatByGroup('tendonK','group',ed)
legLengthString =  tableFormatByGroup('legLength','group',ed)
femurNString =  tableFormatByGroup('femurN','group',ed)
tibiaNString =  tableFormatByGroup('tibiaN','group',ed)
tmtNString =  tableFormatByGroup('tmtN','group',ed)
toeNString =  tableFormatByGroup('toeN','group',ed)


table.Functional = data.frame(totalsEd[,2],massString,bDensityString,perFatString,avMaString,tendonKString, legLengthString, femurNString,tibiaNString,tmtNString,toeNString)
row.names(table.Functional) = c('Restricted','Control','pVal','ES')
colnames(table.Functional) = c('Total animals','mass (kg)','Bone Density cm2','Percent Fat','Moment Arm mm','Tendon Stiffness','Leg Length mm','N.femur Length','n.tibia Length','n.tmt Length','n.toeLength')
table2.Functional = as.data.frame(t(table.Functional))
library(knitr)
pandoc.table(table2.Functional, caption='Table 1: Morpgh data by treatment', justify='center')


```


```{r tableMorph2, echo=FALSE, fig.height=7, fig.width=8, message=FALSE, warning=FALSE, results='asis'}
dta = ed
require('pander')
panderOptions('round',1)
panderOptions('table.style','rmarkdown')
panderOptions("table.split.table", Inf) 
panderOptions('keep.trailing.zeros',TRUE)
library(plyr)
#functional data

lgMassString = tableFormatByGroup('LG','group',dta)
lgAvOflString = tableFormatByGroup('LgAvOfl','group',dta)
fMaxLgString = tableFormatByGroup('fMaxLg','group',dta)
optPenAngLg = tableFormatByGroup('optPenAngLg','group',dta)
tendonSlackLengthLgmm = tableFormatByGroup('tendonSlackLengthLGmm','group',dta)
  
mgMassString = tableFormatByGroup('MG','group',dta)
mgAvOflString = tableFormatByGroup('MgAvOfl','group',dta)
fMaxMgString = tableFormatByGroup('fMaxMg','group',dta)
optPenAngMg = tableFormatByGroup('optPenAngMg','group',dta)
tendonSlackLengthMgmm = tableFormatByGroup('tendonSlackLengthMGmm','group',dta)
   
  
  table.Functional = data.frame(lgMassString,lgAvOflString,fMaxLgString,optPenAngLg,tendonSlackLengthLgmm,mgMassString,mgAvOflString,fMaxMgString,optPenAngMg,tendonSlackLengthMgmm)
row.names(table.Functional) = c('Restricted','Control','pVal','ES')
colnames(table.Functional) = c('LG Mass','Lg OFL','LG F max','LG Pen Ang','LG Tendon SL','MG Mass','Mg OFL','MG F max','MG Pen Ang','MG Tendon SL')
table2.Functional = as.data.frame(t(table.Functional))
library(knitr)
pandoc.table(table2.Functional, caption='Table 1: Morpgh data by treatment', justify='center')


```



## Systems level Question 1:  Can birds restricted from jumping store less energy in their tendon in preparation for a jump??


This is where I make subject specific OSim models and set them all to a range of crouch postures and stimulate LG and MG at 100% (??) to see how much they could strain their tendon.  



```{r elasticstorageLinear, echo=FALSE,fig.height = 5.5, fig.width=6}


cols = c('steelblue4','grey50')
par(mfrow=c(1,2))
par(oma=c(1.5,2,2,0),family='Garamond',xpd=NA,mgp=c(2.2,0.7,0), mar=c(2,3.5,1.5,1))
par(ps = 16,cex=2,cex.main=1.5)
layout(matrix(c(1,1,2,2), nrow = 1, ncol = 4, byrow = TRUE))

ese = ed[ed$group=='E',]
esd = ed[ed$group=='D',]


plot(maxPowerPKg~PEc,data=ed,col='black',pch=21,ylab='Peak Jump Power W/kg',xlab = 'Tendon Strain Energy J',cex=2,bg=c('steelblue4','grey50')[as.numeric(ed$group)],cex.lab = 1.5, cex.axis=1.2 )

plot(workPKg~PEc,data=ed,col='black',pch=21,ylab='Jump Work J/kg',xlab = 'Tendon Strain Energy J',cex=2,bg=c('steelblue4','grey50')[as.numeric(ed$group)],cex.lab = 1.5, cex.axis=1.2 )


esdFit = lm(KE~PE,data=ed)
summary(esdFit)

#abline(esdFit)
t.test(PE~group,data=ed)
```

```{r elasticstorageBoxPlot, echo=FALSE,fig.height = 5.5, fig.width=3}


par(mfrow=c(1,1))
groupNames = c('Restricted','Control')
cols = c('steelblue4','grey50')
boxplot(PE~group,data=ed,col=c('steelblue4','grey50'), names= groupNames, cex.lab = 1.4, cex.axis = 1.2, ylab = 'Energy Storage J', at=c(0.85,1.15),boxwex=0.2, xlim=c(0.7,1.3) )

ese = ed[ed$group=='E',]
esd = ed[ed$group=='D',]

stripchart(ese$PE, vertical=TRUE, add=TRUE, pch=21,bg='grey50',method='jitter',cex=1.5, at=c(1.15),jitter=0.05)
stripchart(esd$PE, vertical=TRUE, add=TRUE, pch=21,bg='steelblue4',method='jitter',cex=1.5, at=c(0.85),jitter=0.05)


```

```{r ModelComparisons, echo=FALSE,results=FALSE,fig.height = 5.5, fig.width=6}
ed = es[es$group=='E' |es$group=='D',]
ed$group = droplevels(ed$group)
ed$avOFL = (ed$LgAvOfl+ed$LgAvOfl)/2
ed$sumMaxF = ed$fMaxLg+ed$fMaxMg
ed$AvLenA0c = ed$LgLenA0c+ed$MgLenA0c
fullModel = lm(PEc~sumMaxF*tendonK*AvLenA0c,data=ed)
AIC(fullModel)
stepResults = stepAIC(fullModel, direction ='both')
summary(stepResults)
AIC(stepResults)

#Individual models
nullModel =  lm(PEc~1,data=ed)
fmaxModel =  lm(PEc~sumMaxF,data=ed)
tendonKModel =  lm(PEc~tendonK,data=ed)
AvLenA0cModel =  lm(PEc~AvLenA0c,data=ed)
summary(nullModel)

summary(fmaxModel)
summary(tendonKModel)
summary(AvLenA0cModel)



aicNull = AIC(nullModel)
aicFmax = AIC(fmaxModel)
aicTendonK = AIC(tendonKModel)
aicLenAo = AIC(AvLenA0cModel)


PredictAIC =c(aicNull,aicFmax,aicTendonK,aicLenAo)
minAIC = min(PredictAIC)
delAIC = PredictAIC-PredictAIC[1]
wAIC = exp(-1/2*delAIC)/exp(-1/2*sum(delAIC))
xtimes =wAIC/wAIC[1]
r2 = 1-(xtimes[2]/xtimes[1])^(2/length(df))

PredictAIC2 =c(-61,aicFmax,aicTendonK,aicLenAo)
minAIC2 = min(PredictAIC2)
delAIC2 = PredictAIC2-min(PredictAIC2)
wAIC2 = exp(-1/2*delAIC2)/exp(-1/2*sum(delAIC2))
xtimes2 =wAIC2[1]/wAIC2
r2 = 1-(xtimes[2]/xtimes[1])^(2/length(df))

```


```{r PETable,echo=FALSE, fig.height=7, fig.width=8, message=FALSE, warning=FALSE, results='asis'}

require('pander')
panderOptions('round',1)
panderOptions('table.style','rmarkdown')
panderOptions("table.split.table", Inf) 
panderOptions('keep.trailing.zeros',TRUE)
library(plyr)


ed = es[es$group=='E' |es$group=='D',]
ed$group = droplevels(ed$group)

#functional data
ed$AvLenA0 = (ed$LgLenA0+ed$MgLenA0)/2
ed$AvLenA100 = (ed$LgLenA100+ed$MgLenA100)/2

ed$AvLenA0c = (ed$LgLenA0c+ed$MgLenA0c)/2
ed$AvLenA100c = (ed$LgLenA100c+ed$MgLenA100c)/2

PeString = tableFormatByGroup('PE','group',ed)
LgFlA0String = tableFormatByGroup('LgLenA0','group',ed)
LgFlA100String = tableFormatByGroup('LgLenA100','group',ed)
MgFlA0String = tableFormatByGroup('MgLenA0','group',ed)
MgFlA100String = tableFormatByGroup('MgLenA100','group',ed)

PecString = tableFormatByGroup('PEc','group',ed)
LgFlA0cString = tableFormatByGroup('LgLenA0c','group',ed)
LgFlA100cString = tableFormatByGroup('LgLenA100c','group',ed)
MgFlA0cString = tableFormatByGroup('MgLenA0c','group',ed)
MgFlA100cString = tableFormatByGroup('MgLenA100c','group',ed)

kneeMaxPeString =  tableFormatByGroup('PEvals_1','group',ed)
ankleMaxPeString =  tableFormatByGroup('PEvals_3','group',ed)

  table3.Functional = data.frame(PeString,LgFlA0String,MgFlA0String,LgFlA100String,MgFlA100String,PecString,LgFlA0cString,MgFlA0cString,LgFlA100cString,MgFlA100cString,kneeMaxPeString,ankleMaxPeString)
row.names(table3.Functional) = c('Restricted','Control','pVal','ES')
colnames(table3.Functional) = c('Energy Storage Potential','LG FL at A0','MG FL at A0','LG FL at A100','MG FL at A100','Energy Storage Potential','LG FL at A0','MG FL at A0','LG FL at A100','MG FL at A100','Knee Angle at Max PE','Ankle Angle at Max PE')
table3.Functional = as.data.frame(t(table3.Functional))
library(knitr)
pandoc.table(table3.Functional, caption='Table 3: Energy Storage By Group', justify='center')



```

